---
title: "Some Quarto Tips for Academic Presentation"
date: 2022-12-28
author: Kazuharu Yanagimoto
draft: false
categories: [R, Quarto]
image: https://source.unsplash.com/F2KRf_QfCqw
execute:
    eval: false
---

## Are You a Beamer Fan?
The other day, I used slides made with [Quarto](https://quarto.org/) in a presentation at a conference.
In the Economics community, 98% of people use Beamer of $\LaTeX$,
and about 2% use PowerPoint for their presentations.
I personally don't like Beamer's design,
and I think PowerPoint has a problem with maintainability
(i.e., can it be easily changed) in case of ongoing projects.
I like Quarto for the following reasons:

1. **Reasonably good design**. And highly customizable.
1. **Time saving**. Slides are written in the markdown notation, so you can make it with the same amount of effort as Beamer (unless you care about the design😉)
1. **Easily maintainable**. Variables can be used in the text, so the numbers are automatically updated even if the model or analysis is slightly changed.

Here, I write some tips on Quarto for academic presentations. For the basic slide creation in Quarto, please refer to the [official documentation]((https://quarto.org/docs/presentations/revealjs/))
or Tom Mock's [slides]((https://rstudio-conf-2022.github.io/get-started-quarto/materials/05-presentations.html#/presentations)) at the RStudio conference.
The slides used at the conference presentation are not yet ready for public release,
so I prepare these example slides

```{=html}
<iframe class="slide-deck" src="https://kazuyanagimoto.com/quarto-slides-example/code/slides/quarto_academic_tips/slides.html"></iframe>
```


[ {{< fa chalkboard-teacher >}} Slides](https://kazuyanagimoto.com/quarto-slides-example/code/slides/quarto_academic_tips/slides.html) &nbsp; 
[ {{< fa file-pdf >}} PDF](https://kazuyanagimoto.com/quarto-slides-example/code/slides/quarto_academic_tips/slides.pdf) &nbsp; 
[ {{< fa brands github >}} Code](https://github.com/nicetak/quarto-slides-example)


## Data

I render tables and plots inside of slides.
In other words, I am not using the image files for the article.
This is because

1. Proper font sizes in the article are **too small** in presentation slides
1. **Highlighting** is not easy. You need to create another figure/table or use complicated TikZ
1. I want to **unify the design** of slides and figures&tables

Hence, I create and save a data frame in a separate file and load it just before plotting it.

```{.r}
load("tb_hoge.rds")
tb_hoge |>
  ggplot(aes(x, y)) +
  geom_point()
```

You can use this data frame to refer to its values in a slide text.

```{markdown}
In 2021, `r tb_hoge$value[tb_hoge$year == 2021]`% of people…
```

This is convenient in case you change your analysis slightly and the values in `tb_hoge.rds` change.^[
  In economics, changing the age group of the sample or the parameters of the model
  sometimes lead to changes in the results.
  In addition, we sometimes found mistakes in cleaning the data.
]
The values will be updated by re-rendering slides.


## Theme

There are 11 built-in themes provided for Reveal presentations.
These are good designs, but to match them with figures&tables,
I write `custom.scss` and make my own theme.
Currently, it is like [xaringanthemer](https://pkg.garrickadenbuie.com/xaringanthemer/)'s
`style_mono_accent()` theme.
You can check about the customizable parts in the [official document](https://quarto.org/docs/presentations/revealjs/themes.html#customizing-themes).

```{css}
#| code-fold: true

/*-- scss:defaults --*/

@import url(https://fonts.googleapis.com/css?family=Montserrat:300,300i&display=swap);
@import url(https://fonts.googleapis.com/css?family=Josefin+Sans&display=swap);
@import url(https://fonts.googleapis.com/css?family=Fira+Mono&display=swap);

// fonts
$font-family-sans-serif: Montserrat, sans-serif !default;
$font-family-monospace: "Fira Mono", monospace  !default;

// colors
$body-bg: #fff !default;
$body-color: #272822 !default;
$link-color: #055099 !default;

// headings
$presentation-heading-font: "Josefin Sans", sans-serif !default;
$presentation-heading-color: #1C5253 !default;

$h1-font-size: 1.6em !default;
$h2-font-size: 1.3em !default;
$h3-font-size: 1.15em !default;
$h4-font-size: 1em !default;

// code blocks
$code-block-border-color: #93a1a1 !default;
$code-color: $presentation-heading-color;

//
// Color system
//

// stylelint-disable
$white:    #fff !default;
$gray-100: #fdf6e3 !default;
$gray-200: #eee8d5 !default;
$gray-300: #dee2e6 !default;
$gray-400: #ced4da !default;
$gray-500: #adb5bd !default;
$gray-600: #839496 !default;
$gray-700: #495057 !default;
$gray-800: #073642 !default;
$gray-900: #002b36 !default;
$black:    #000 !default;

/*-- scss:rules --*/

.title{
    color: $white;
    font-size: 2.5em;
    font-family: $presentation-heading-font;
    position: absolute;
    top: 20%
}

.subtitle{
    color: $gray-100;
    font-size: $h2-font-size;
    font-family: $presentation-heading-font;
    position: absolute;
    top: 33%
}


.author {
    color: $white;
    font-size: $h3-font-size;
    font-family: $presentation-heading-font;
    position: absolute;
    bottom: 10%;
}

.date {
    color: $white;
    font-size: $h4-font-size;
    font-family: $presentation-heading-font;
    position: absolute;
    bottom: 5%;
}


.message{
    color: $presentation-heading-color;
    font-size: $h2-font-size;
    font-family: $presentation-heading-font;
    text-align: center;
}

.space-left {
    margin-left: 5%;
}


.center-xy {
    margin: 0;
    position: absolute;
    top: 50%;
    left: 50%;
    -ms-transform: translateY(-50%), translateX(-50%);
    transform: translateY(-50%), translateX(-50%);
}


.small {
    font-size: 0.8em
}

.Large {
    font-size: 1.6em
}

section.has-dark-background a {
    color: $white;
    border-style: dashed;
    border-bottom-width: 1px;
}

section.has-dark-background a:hover {
    color: $white;
    border-bottom: solid
}

```


### Ad-hoc CSS

I write frequently used CSS settings in `custom.css`,
but sometimes I want to use a CSS environment on an ad-hoc basis
(for example, to change the font size slightly.)

In these cases, you can write them in `style=` environment:

```{.md}
::: {style="font-size: 0.68"}

Sentences you want to make smaller

:::
```

### Fragments

Quarto (or Reveal.js) allows some complicated animation.
If you are interested, you might want to see Tom Mock's
[this part]((https://rstudio-conf-2022.github.io/get-started-quarto/materials/05-presentations.html#/lists))
of the slides.

Apart from this, as discussed
[here](https://community.rstudio.com/t/quarto-revealjs-presentation-2-columns-with-pause/151950),
the original fragment separator `. . .` does not correctly work in environments like `columns`.
From my experience, `fragment` environment solves them.

## Plots
### Theme

Minimizing elements is a starting point of slide design.
For that purpose, it is important to use the same fonts and colors for figures (and tables)
as the slides.
This is the reason why I used to use [xaringanthemer]((https://pkg.garrickadenbuie.com/xaringanthemer/)),
which automatically generates the `ggplot2` theme corresponding to the slides' theme.
Unfortunately, there have been no packages for Quarto slides (as far as I know),
I declare this `theme_quarto()` function.

```{r}
#| code-fold: true
blend_colors <- function(x, y, alpha = 0.5) {
  x <- colorspace::hex2RGB(x)
  y <- colorspace::hex2RGB(y)
  z <- colorspace::mixcolor(alpha, x, y)
  colorspace::hex(z)
}
color_blender <- function(x, y) function(alpha = 0.5) blend_colors(x, y, alpha)
theme_quarto <- function (
    text_color = color_text,
    background_color = color_bg,
    text_font = font_text,
    text_font_size = 30,
    accent_color = color_base,
    title_font = font_title,
    title_font_size = 30
){
    blend <- color_blender(text_color, background_color)
    ggplot2::theme(
        line = ggplot2::element_line(color = blend(0.2)),
        rect = ggplot2::element_rect(fill = background_color),
        title = ggplot2::element_text(
            color = accent_color,
            family = title_font,
            size = title_font_size
        ),
        plot.background = ggplot2::element_rect(
            fill = background_color,
            color = background_color
        ),
        panel.background = ggplot2::element_rect(
            fill = background_color,
            color = background_color
        ),
        panel.grid.major = ggplot2::element_line(
            color = blend(0.8),
            inherit.blank = TRUE
        ),
        panel.grid.minor = ggplot2::element_line(
            color = blend(0.9),
            inherit.blank = TRUE
        ),
        axis.title = ggplot2::element_text(size = title_font_size * 0.8),
        axis.ticks = ggplot2::element_line(color = blend(0.8)),
        axis.text = ggplot2::element_text(color = blend(0.4), size = title_font_size * 0.7),
        legend.key = ggplot2::element_rect(fill = "transparent", colour = NA),
        legend.text = ggplot2::element_text(size = title_font_size * 0.8, family = title_font),
        plot.caption = ggplot2::element_text(
            size = text_font_size * 0.8,
            color = blend(0.3)
        )
    )
}
```

While the original `xaringanthemer::theme_xaringan()` redefines
the primary color of `ggplot2::geom_*` by `ggplot2::update_geom_defaults`,
my function does not.
I did not overwrite it because color palettes must be specified manually for 2+ color cases, anyway.
You can use it like:

```{r}
color_base <- "#1C5253"
tb_hoge |>
  ggplot(aes(x, y)) +
  geom_point(color = color_base) +
  theme_quarto()
```

### Highlighting

The most famous way to highlight plots should be [gghighlight](https://cran.r-project.org/web/packages/gghighlight/vignettes/gghighlight.html).
You can proceed with your stories smoothly and clearly by

```{r filename = "First Slide"}
p <- tb_hoge |>
  ggplot(aes(x, y)) +
  geom_col(color = color_base) +
  theme_quarto()

p
```

```{r filename = "Second Slide"}
p + gghighlight(x == "Japan")
```


## Tables

### `markdown` vs. `kableExtra` vs. `gt`?

I think they are three popular ways to create tables in HTML slides.
while each of them has pros and cons, I am using `gt` because

**Markdown Table**

- No highlighting
- No `multirow` nor `multicol` from $\LaTeX$

**kableExtra**

- Syntax is a little more complicated than `gt` (e.g., need to mutate a string column for percentages.)
- Highlighting is slightly more complicated than `gt`

Note that I use `kableExtra` for tables in an article due to the following problem.

### Problems with Mathematic Expressions

Unfortunately, it is difficult to use mathematical expressions in a table in Quarto.
`gt` does not support $\LaTeX$ notation.^[
  This [issue](https://github.com/rstudio/gt/issues/375) has been raised in 2019,
  but it does not seem to be solved quickly.
] Also, mathematical expressions of `kableExtra` are not rendered in Quarto ([GitHub Issue](https://github.com/quarto-dev/quarto-cli/issues/555)).
You can solve it by

```{.r}
kableExtra::kbl(data, format="markdown")
```

but it contains the abovementioned disadvantages of markdown tables.

Then, I compromise by using

- Unicode Greek letters (θ, τ, δ, etc.)
- HTML tags (`<sup>` and `<sub>`)

and evaluating them by `gt::fmt_markdown()`.
Mathematical expressions in tables are usually simple,
so I don't have any problems so far.

```{r}
tibble(a = c("θ = 0", "τ<sup>δ-1</sup>")) |>
  gt() |>
  fmt_markdown(columns = everything())
```

### Highlighting

The gt package itself can also highlight cells in a table with the function `tab_style()`.
However, this requires redundant writing,
so `gtExtras::gt_highlight_rows()` and `gtExtras::gt_highlight_cols()` are useful
if you want to highlight row/column-wise.

## Export

### HTML or PDF?

Although I think you use a prepared PC for your presentation,
since it is difficult to imagine a PC without a browser,
the HTML format is also acceptable.
If you write the following in the YAML header,
only one self-contained HTML file will be output.
So you can easily carry it around like a PDF file.^[
  The option `self-contained: true` produces a deprecated warning.
]

```{.yaml}
format:
  revealjs:
    standalone: true
    embed-resources: true
```

However, you will be asked to submit your slides in PDF format in various situations.
And whether or not you actually present your presentation in PDF format,
it is a good idea to have PDF slides on hand for insurance purposes.

1. Open the HTML file in your browser and press <kbd>E</kbd>
1. Click "Print" on the browser, then "Save as PDF"

If you want to separate fragments, specify `pdf-separate-fragments: true` in YAML.

### Problems with Hyperlink

A hyperlink that jumps to each slide is essentially a reference to an `<h2>` tag,
so the link can be written as:

```{.markdown}
## A Slide {#sec-slide}

<a href="#/sec-detail">detail↗</a>

## A Detail Slide {#sec-detail}

<a href="#/sec-slide">↩︎</a>
```

However, as far as I know, the link is not reflected in the above PDF output. If anyone knows a solution to this problem, I would appreciate it if you could let me know.

## Fin.
These are some of my ideas for using Quarto for academic presentations.
I think that Quarto's slides are quite suitable for academic presentations,
because of their simple syntax, high degree of freedom in design,
and high maintainability in numbers.
I hope this is useful for you.
